\documentclass[a4paper, 11pt]{article}

%\title{Unused Title}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{tabulary}
\usepackage{adjustbox}
\usepackage{blindtext}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{csquotes}
\usepackage[T1]{fontenc}

% Use helvet if uarial cannot be installed
%\usepackage{uarial}
\usepackage[scaled]{helvet}

\renewcommand{\familydefault}{\sfdefault}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{courier}
\usepackage{setspace}
\usepackage[table,svgnames]{xcolor}
\usepackage{fancyvrb} 
\usepackage{listings}
\usepackage{caption}
\usepackage{longtable}
\usepackage{relsize}
\usepackage{tfrupee}
\usepackage{rotating}
\usepackage{lipsum}
\usepackage{subcaption}
\usepackage{float}
\usepackage{aliascnt}
\usepackage{svg}
\usepackage{placeins}
\usepackage{lineno}
\linenumbers
\usepackage{cleveref}
\captionsetup[table]{labelfont={small,bf,it},textfont={small,it}}
\captionsetup[figure]{labelfont={small,bf,it},textfont={small,it}}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\usepackage[style=authoryear, backend=bibtex]{biblatex}
\bibliography{report.bib}

\makeatletter
\newcommand\footnoteref[1]{\protected@xdef\@thefnmark{\ref{#1}}\@footnotemark}
\makeatother

\newaliascnt{eqfloat}{equation}
\newfloat{eqfloat}{h}{eqflts}
\floatname{eqfloat}{Equation}

\newcommand*{\ORGeqfloat}{}
\let\ORGeqfloat\eqfloat
\def\eqfloat{%
	\let\ORIGINALcaption\caption
	\def\caption{%
		\addtocounter{equation}{-1}%
		\ORIGINALcaption
	}%
	\ORGeqfloat
}

\addto\captionsenglish{% Replace "english" with the language you use
	\renewcommand{\contentsname}%
	{List of Contents}%
}

\newcommand\tab[1][1cm]{\hspace*{#1}}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
	backgroundcolor=\color{backcolour},   
	commentstyle=\color{codegreen},
	keywordstyle=\color{magenta},
	numberstyle=\tiny\color{codegray},
	stringstyle=\color{codepurple},
	basicstyle=\ttfamily\footnotesize,
	breakatwhitespace=false,         
	breaklines=true,                 
	captionpos=b,                    
	keepspaces=true,                 
	numbers=left,                    
	numbersep=5pt,                  
	showspaces=false,                
	showstringspaces=false,
	showtabs=false,                  
	tabsize=2,
	xleftmargin=0.5cm,
	xrightmargin=-0.8cm,
	frame=lr,
	%	framesep=-5pt,
	framerule=0pt
}

\lstset{style=mystyle}

\definecolor{Teal}{RGB}{0,128,128}
\definecolor{NewBlue1}{RGB}{4,100,226}
\definecolor{NiceBlue}{RGB}{63,104,132}
\definecolor{DarkRed}{RGB}{14,53,59}
\definecolor{NewBlue2}{RGB}{62,100,125}
\definecolor{NewBlue3}{RGB}{44,100,128}

\hypersetup{
	colorlinks,
	citecolor=NiceBlue,
	linkcolor=NewBlue1,
	urlcolor=Blue
	%	citebordercolor=Violet,
	%	filebordercolor=Red,
	%	linkbordercolor=Blue
}

\usepackage{geometry}
\linespread{1.25}
\usepackage[parfill]{parskip} % Avoid indentation

\geometry{
	a4paper,
	left=4cm,
	right=2.5cm,
	top=2.5cm,
	bottom=2.5cm,
}


\begin{document}
	\pagenumbering{gobble}
	\begin{center}
		{\large IMPERIAL COLLEGE LONDON}
	\end{center}
	%	\maketitle
	\vspace{6cm}
	
	\begin{center}
		
		\Huge Themes in IUCN Red List assessments revealed by Machine Reading\\		
		\vspace{.5cm}		
		\large {Word Count: 4,735}
		
	\end{center}
	\vspace{2.5cm}
	\begin{center}
		\Large Yew Shen Lim
	\end{center}
	
	\begin{center}
		{\large August 2021}
	\end{center}

	\vspace{7cm}
	\begin{center}
		{\large A thesis submitted in partial fulfilment of the \\requirements for the Master of Research at Imperial College London \\ Sumitted for the MRes in Computational Methods in Ecology and Evolution degree}
	\end{center}		

	\newpage
	\pagenumbering{Roman}

\section*{Declaration}

Both the text data and mammals range data used were downloaded from the IUCN Red List website.

All of the code and graphics were written and created in Python and R. I was responsible for preprocessing the text and conducting the analysis.

My supervisor Dr James Rosindell provided guidance for my project's development in weekly meetings.

\pagebreak

% \section*{Acknowledgments}
% %\include{Acknowledgments.tex}
% \blindtext
% \pagebreak

% \section*{Notation}
% The extent to which you list any symbols used in your report must be left to your discretion. Symbols, which are used in several parts of your report, should preferably be listed before the main text for easy reference. Symbols which are used only once or in one part of the report may be referred to in that part only. Generally, try to place yourself in the position of a reader with average background knowledge and arrange the notation in a manner which will be most convenient for him/her to follow.
% \pagebreak

%\section*{List of Contents}
\tableofcontents
\pagebreak
%\listoffigures
%\listoftables
%\pagebreak
\newpage
%\section{Notation}
%\textbf{IF APPLICABLE}
%%\include{Acknowledgments.tex}


\cleardoublepage\pagenumbering{arabic}


\section*{Abstract}
%\textbf{MAXIMUM 250 WORDS}
The IUCN Red List is the primary source of extinction risk for life on Earth. The availability of the IUCN assessment justifications in plain text format allows for the application of text mining techniques to discover themes and identify potential gaps in existing literature. This study is the first to apply text mining on IUCN assessments. I applied topic modelling on the texts with the aim to discover latent themes. I developed a model using the texts, and the model identified 17 different topics from the texts which were then analysed from different angles. The results of this study complements the literature surrounding biodiversity loss and conservation efforts. This study also seeks to fill the gap between how the Red List is used and how it could be used, laying a foundation for further downstream analyses.
\pagebreak

\section{Introduction}
\input{introduction.tex}
\pagebreak

\section{Methods}
\input{methods.tex}
\pagebreak

\section{Results}
\input{results.tex}
\pagebreak

\section{Discussion}
\input{discussion.tex}
\pagebreak

\section{Conclusion}
\input{conclusion.tex}
\pagebreak

\section{References}
\printbibliography[heading=none]
\pagebreak

\pagenumbering{roman}

\section{Appendix}
\input{appendix.tex}

\end{document}   